---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# learningr

<!-- badges: start -->
[![binder](https://github.com/mahendra-mariadassou/learningr/workflows/binder/badge.svg)](https://mybinder.org/v2/gh/mahendra-mariadassou/learningr/master)
<!-- badges: end -->

The goal of learningr is to provide interactive tutorials for the statistics classes given by Tristan for UE 2.3. The tutorial are shamelessly taken from [rstudio-education](https://rstudio.cloud/learn/primers) for the basics of R and the tidyverse. Tutorials are packaged so you can install them on you computer and do the exercices without access to an internet connection. 

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("mahendra-mariadassou/learningr")
```

<!-- If the installation of `devtools` fails, you may need to install some system dependencies (for example on ubuntu/debian): -->
<!-- ```{bash, eval = FALSE} -->
<!-- sudo apt-get install libxml2-dev libcurl4-openssl-dev libssl-dev -->
<!-- ``` -->

<!-- On Mac, you may need some of the compiling tools listed [here](https://cran.r-project.org/bin/macosx/tools/) -->

On **Windows**: you may need **Rtools** and **git** 

- **Rtools**: visit the dedicated [page](https://cran.r-project.org/bin/windows/Rtools/), download the suggested exe and install it on your computer
- **git**: visit the dedicated [page](https://git-scm.com/download/win), download the suggested exe and install it on your computer

Alternatively, you can use create a remote R session to complete the tutorial by launching binder:  [![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/mahendra-mariadassou/learningr/master?urlpath=rstudio)

You only need a web browser, no account or anything. The main drawback of this solution (compared to the previous ones) is that you lose your progress each time you launch a new session. 

## Starting a tutorial

This package is intended for use with `learnr`:

```{r example, eval = FALSE}
library(learnr)
```

### Basics of programmings 

```{r programming, eval = FALSE}
## Launch only one tutorial at the time!!
learnr::run_tutorial("01_programming_basics", package = "learningr")
learnr::run_tutorial("03_visualisation_basics", package = "learningr")
```

### R and stats

```{r stats, eval = FALSE}
learnr::run_tutorial("02_random_variables", package = "learningr")
```

### Data visualization

```{r dataviz, eval = FALSE}
learnr::run_tutorial("04.1_exploratory_data_analysis", package = "learningr")
learnr::run_tutorial("04.2_barcharts", package = "learningr")
learnr::run_tutorial("04.3_histograms", package = "learningr")
learnr::run_tutorial("04.4_boxplots", package = "learningr")
learnr::run_tutorial("04.5_scatterplots", package = "learningr")
learnr::run_tutorial("04.6_linegraphs", package = "learningr")
```

## Data manipulation

```{r datamanipulation, eval = FALSE}
learnr::run_tutorial("05.1_tibbles", package = "learningr")
learnr::run_tutorial("05.2_isolating_data", package = "learningr")
learnr::run_tutorial("05.3_summaries", package = "learningr")
```

### Estimation and confidence intervals

```{r confint, eval = FALSE}
learnr::run_tutorial("06.1_sampling", package = "learningr")
learnr::run_tutorial("06.2_testing", package = "learningr")
```

