---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# learningr

<!-- badges: start -->
<!-- badges: end -->

The goal of learningr is to serve a companion tutorials for the statistics classes given by Tristan for UE 2.4

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mahendra-mariadassou/learningr")
```

If the installation of `devtools` fails, you may need to install some system dependencies (for example on ubuntu/debian):
```{bash, eval = FALSE}
sudo apt-get install libxml2-dev libcurl4-openssl-dev libssl-dev
```

On Mac, you may need some of the compiling tools listed [here](https://cran.r-project.org/bin/macosx/tools/)

On **Windows**: you'll need **Rtools** and **git** 

- **Rtools**: visit the dedicated [page](https://cran.r-project.org/bin/windows/Rtools/), download the suggested exe and install it on your computer
- **git**: visit the dedicated [page](https://git-scm.com/download/win), download the suggested exe and install it on your computer

## Example

This package is intended for use with `learnr`:

```{r example, eval = FALSE}
library(learnr)
learnr::run_tutorial("01_programming_basics", package = "learningr")
```
